---
title: 'Probs'
subtitle: 'scratch'
author: 'fnaufel'
email: 'https://fnaufel.github.io/'
date: 'Versão de `r format(Sys.Date(), "%d/%m/%Y")`'
lang: 'pt'
# TODO: include LaTeX fields
output: 
  html_document:
    theme: readable
           # https://bootswatch.com/3/readable/
    highlight: tango
    css: styles.css
    toc: true
    toc_depth: 2
    number_sections: true
    fig_caption: true
    df_print: paged
    self_contained: true
    # code_folding: show
    # keep_md: true
    # includes:
    #   in_header: header.html
    #   before_body: doc_prefix.html
    #   after_body: doc_suffix.html
---

```{r setup, include=FALSE}
library(knitr)

opts_chunk$set(
  echo = FALSE, 
  # collapse = TRUE,
  # cache = TRUE,
  out.width = "75%",
  fig.align = 'center',
  fig.width = 6,
  fig.show = "hold"
)

options(dplyr.print_min = 6, dplyr.print_max = 6)

# Supress crayon output
options(crayon.enabled = FALSE)

# For nice dataframe summaries
library(summarytools)
st_options(
  plain.ascii = FALSE,
  dfSummary.varnumbers = FALSE,
  dfSummary.style = 'grid',
  dfSummary.graph.magnif = .75
)

# Tidy!
library(tidyverse)

# Sober theme for ggplot
theme_set(
  theme_linedraw() +                         # Set simple theme for ggplot
    theme(                                   # with some tweaks
      axis.title.y.left = element_text(
         angle = 0,                          # Never rotate y axis title
         margin = margin(r = 20),            # Separate y axis title a little 
         vjust = .5                          # Leave y axis title in the middle
      ),
      axis.title.x.bottom = element_text(
         margin = margin(t = 20)             # Separate x axis title a little 
      ),
      axis.line = element_blank(),           # No axis lines
      panel.border = element_blank(),        # No frame
      panel.grid.minor = element_blank()     # No grid minor lines
    )
)

# Avoid scientific notation and use a comma as decimal separator
options(
  scipen = 15,
  OutDec = ','
)

# Format a number with thousand separators (default point)
# and decimal comma enclosed in curly braces for LaTeX printing.
# CAREFUL: if called outside math mode, will print the braces!
fm <- function(x, big = '.', decimal = '{,}') {
  if (!is.numeric(x)) {
    x
  } else {
    prettyNum(x, big.mark = big, decimal.mark = decimal)
  }

}

# To center the results of a chunk (image, video etc.)
# Usage: 
#         out.extra=center()
#         
center <- function(){
  
  if (is_html_output()) {
    'class="center"'
  }
  
}


# To embed YT videos in HTML and the link (centered) in LaTeX
embed_yt <- function(code) {

  if (is_html_output()) {
    include_url(
      paste0(
        'https://www.youtube.com/embed/',
        code
      )
    )
  } else {
    cat(
      paste0(
        '```{=latex}\n',
        '\\begin{center} \\url{https://youtu.be/',
        code,
        '} \\end{center}\n',
        '```'
      )
    )
  }
  
}

source('../packages.R')

loadd(
  df,
  df_sucessos_raras, 
  n_raras,
  df_sucessos_freqs,
  df_vetor,
  df_summary,
  df_regress,
  arrec_resultados, 
  n_recentes, 
  rsq_CEF,
  quadra_resultados,
  quina_resultados,
  sena_resultados,
  premios
)

opts_knit$set(root.dir = '/home/fnaufel/Documents/DataAnalysis/Megasena')

```

```{js javascript-init, echo=FALSE}

// Make off-site links open in a new window/tab
function changeTargets() {
  $("a").attr(
    "target", function() {
      // Load local links locally
      if (this.host == location.host) return "_self"
      // Load off-site links in a new window
      else return "_blank";
    }
  );
}

// Execute when document is ready
$(
  changeTargets
)
```



# Probabilidades

A fórmula geral para 

* $a$ acertos
* $j$ números jogados
* $s$ números sorteados dentre
* $t$ números no total

é

$$
\Pr(a \mid j, s, t) = \frac{{j \choose a}{t - j \choose s - a}}{t \choose s}
$$

O total de resultados possíveis é $t \choose s$.

A idéia é que, fixados os $j$ números jogados, para determinar a quantidade de resultados que têm exatamente $a$ acertos, é preciso

* determinar os acertos dentre os jogados: $j \choose a$

* determinar os números sorteados restantes (que não podem ter sido jogados): $t - j \choose s - a$

Na verdade, o número de acertos segue a [distribuição hipergeométrica](https://pt.wikipedia.org/wiki/Distribui%C3%A7%C3%A3o_hipergeom%C3%A9trica).

```{r}
total <- 6
sorteados <- 3
jogados <- sorteados:(total-1)

acertos_3 <- dhyper(x = 3, m = jogados, n = total - jogados, k = sorteados)
acertos_2 <- dhyper(x = 2, m = jogados, n = total - jogados, k = sorteados)

chances_3 <- (1/acertos_3) %>% ceiling()
chances_2 <- (1/acertos_2) %>% ceiling()

probs <- tibble(jogados, acertos_3, acertos_2)
chances <- tibble(jogados, chances_3, chances_2)

probs
chances
```


```{r}
total <- 10
sorteados <- 3
jogados <- sorteados:(min(sorteados + 9, total - 1))
acertos <- sorteados:(max(sorteados - 2, 1))
names(acertos) <- acertos

probs <- acertos %>%
  map_dfc(dhyper, m = jogados, n = total - jogados, k = sorteados) %>% 
  mutate(jogados = jogados, .before = 1)

probs %>% 
  kbl(format.args = list(big.mark = '.'), digits = 9) %>% 
  kable_paper(
    c('striped', 'hover'),
    full_width = FALSE
  )

chances <- probs %>% 
  mutate(across(!1, ~ round(1/.x, 0)))

chances %>% 
  kbl(format.args = list(big.mark = '.')) %>% 
  kable_paper(
    c('striped', 'hover'),
    full_width = FALSE
  )

```
